---
title: "Insey Winsey Spider: Monte Carlo simulation and theorectical explanation"
author: "Jitao david Zhang"
date: "21 Juli 2018"
output: html_document
---

# Background 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
shapeDice <- function() sample(1:6, 1)
weatherDice <- function() sample(c("sun", "rain"), 1)
```

# Game logic

```{r logic}
SHAPE_CODES <- c("START"=0,
  "half-circle"=1,
  "diamond"=2,
  "square"=3,
  "rect"=4,
  "circle"=5,
  "triangle"=6)

YELLOW_SHAPES <- c("START", "half-circle", "diamond", "square", "rect",
            "circle", "triangle", "half-circle", "diamond", "square", "rect")
YELLOW_CODES <- SHAPE_CODES [YELLOW_SHAPES]

game <- function(seed=1, codes=YELLOW_CODES, debug=TRUE) {
  set.seed(seed)
  step <- 0
  pos <- 1
  goal <- length(codes)
  while(pos != goal) {
    if(debug) {
      oldPos <- pos
    }
    codesAhead <- codes[(pos+1):goal]
    newShape <- shapeDice()
    step <- step+1
    if(newShape %in% codesAhead) {
      pos <- pos + match(newShape, codesAhead)
    }
    weather <- weatherDice()
    if(weather == "rain") {
      pos <- 1
    }
    if(debug) {
      cat(sprintf("old pos: %d; shape:%s; weather:%s; new pos:%d\n", oldPos, names(SHAPE_CODES)[match(newShape, SHAPE_CODES)], weather, pos))
    }
  }
  return(step)
}
```

# An example

Here is a simulated run of the game.

```{r}
game(seed=1, debug=TRUE)
```

I manually checked three seeds (1, 2, 1887), and the logic seems to be correct.

# Monte Carlo simulation
We use Monte Carlo simulation to figure out the distribution of steps.

```{r monteCarlo}
mcN <- 20000
system.time(mcRes <- sapply(seq(1, mcN), game, debug=FALSE))
```

By simulation, we found that it takes a median of 26 steps to finish a game. Suppose that each step of a gamer costs 20 seconds (throw the dice, move the spider, tell the weather, and celebrate or 'splash'), it takes a median of 8.5 min for her or him to win. That means, the median game time is probably around 17 minutes. 

Following the calculating, the interquantile of the game time is between 8 min and 33 min.

```{r summ}
summary(mcRes)
```

Next, we visualize the results of Monte Carlo simulation with a histogram.

```{r hist}
hist(mcRes, breaks=200, col="lightblue", ylab="Count", main="Steps of a gamer by Monte Carlo simulation")
```

# Distribution fitting and theorectical explanations

The last question is, what distribution does this follow? It looks like an expotential distribution. The empirical fit seems very good.

```{r dist, fig.height=5, fig.width=7}
beta <- mean(mcRes-1) ## 1/lambda
beta2 <- var(mcRes-1) ## 1/lambda^2
beta2/beta^2 ## near 1
lambda <- 1/beta
{
  hist(mcRes, breaks=200, col="lightblue", ylab="Probability", main="Monte Carlo simulation and distribution fitting", 
       prob=TRUE)
  lines(1:300+1,
        dexp(1:300, rate=lambda), col="red", lwd=2.5)
  legend("topright", pch=c(22, NA), col=c("black", "red"), lty=c(NA, 1), 
         pt.bg=c("lightblue", NA),
         pt.cex = c(1.5, NA),
         legend=c("Monte Carlo", "Expotential distribution"))
}
```

Now comes the last question: can we explain the observed good fit of the expotential curve?

